### 2012-09-16 doc/prelim/document/chapters/linear_problem/linear_problem.tex
added section on preconditioning stochastic methods

\section{Preconditioning the Linear System}
\label{sec:linear_preconditioning}

>>>>

\subsubsection{Preconditioning Stochastic Methods}
\label{subsubsec:stochastic_preconditioning}
In most cases, at least a minimal amount of \textit{preconditioning}
of the linear system will be required in order to use the class of
stochastic methods described. Although these methods have no symmetry
requirements for convergence, they do require that the spectral radius
of the iteration matrix be less than one. To achieve this, a Jacobi
precondtioner, a form of left preconditioning, is used such that the
preconditioning matrix $\ve{D}$ is:
\begin{equation}
  \ve{M} = diag(\ve{A})\:,
  \label{eq:jacobi_preconditioner}
\end{equation}
such that its application means we are instead solving the following
linear system:
\begin{equation}
  \ve{M}^{-1}\ve{A}\ve{x} = \ve{M}^{-1}\ve{b}\:.
  \label{eq:jacobi_precond_linear_problem}
\end{equation}
Next, we can apply MCSA to solve
Eq~(\ref{eq:jacobi_precond_linear_problem}): 
\begin{subequations}
  \begin{gather}
    \ve{x}^{k+1/2} = (\ve{I} - \ve{M}^{-1}\ve{A})\ve{x}^k +
    \ve{M}^{-1}\ve{b}\:,\\ \ve{r}^{k+1/2} = \ve{M}^{-1}\ve{b} -
    \ve{M}^{-1}\ve{A}\ve{x}^{k+1/2}\:,\\ \ve{M}^{-1}\ve{A}\delta\ve{x}^{k+1/2}
    = \ve{r}^{k+1/2}\:,\\ \ve{x}^{k+1} = \ve{x}^{k+1/2} + \delta
    \ve{x}^{k+1/2}\:.
    \label{eq:jacobi_preconditioned_mcsa}
  \end{gather}
\end{subequations}
Choosing Jacobi preconditioning with MCSA is advantageous for several
reasons. First, $\rho(\ve{I} - \ve{M}^{-1}\ve{A}) < 1$ is true for
all $\ve{A}$ and is easy to formulate as the inversion of $\ve{M}$ is
trivial. Second, because the adjoint Monte Carlo method used within MCSA to
compute the correction operates on a linear problem with the
preconditioned operator, then $\ve{H}$ in the adjoint solver will have
a zero term in each of its diagonal elements, thereby eliminating
all in-state transitions during the random walk sequence. Because of
this, Jacobi preconditioning should always be performed, regardless of
any other preconditioning that is applied to the system.
